### 2024-09-20 paper/paper.qmd
create some graph making files

```{python}
#| echo: false
# Execute the external Python script
with open(f"../code/model_generation/dow_bar.py") as script_file:
    exec(script_file.read())
```

```{python}
#| echo: false
# Execute the external Python script
with open("../code/model_generation/individual_beach_scatter.py") as script_file:
    exec(script_file.read())
```

```{python}
import pandas as pd
import matplotlib.pyplot as plt
import seaborn as sns
from matplotlib.ticker import MaxNLocator

# Sample data creation (replace this with your actual data)
data = pd.read_csv("../data/cleaned_water_quality_data.csv")

# Create a DataFrame
df = pd.DataFrame(data)
# Add a column for the day of the week
df['day_of_week'] = df['collectiondate'].dt.day_name()

# Days of the week order for plotting
days_of_week = ['Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday', 'Saturday', 'Sunday']

# Create scatter plots for each day of the week
for day in days_of_week:
    # Filter data for the specific day of the week
    day_df = df[df['day_of_week'] == day]

    # Create the scatter plot
    plt.figure(figsize=(14, 7))
    sns.scatterplot(data=day_df, x='collectiondate', y='ecoli', hue='beachname', palette='Set1', s=75, alpha=0.6)
    plt.title(f'E. coli Levels Over Time - {day}', fontsize=16)
    plt.xlabel('Collection Date', fontsize=14)
    plt.ylabel('E. coli Level', fontsize=14)
    plt.gca().xaxis.set_major_locator(MaxNLocator(nbins=8))
    plt.gca().xaxis.set_major_formatter(plt.matplotlib.dates.DateFormatter('%b %Y'))
    plt.xticks(rotation=45, ha='right')
      # Log scale for better representation
    plt.grid(visible=True, linestyle='--', linewidth=0.5, alpha=0.7)
    plt.legend(title='Beach Name', loc='upper left', bbox_to_anchor=(1.05, 1))
    plt.tight_layout()
    plt.show()

```

### 2024-09-20 code/model_generation/dow_bar.py

@@ -1,0 +1,35 @@
+import pandas as pd
+import matplotlib.pyplot as plt
+import seaborn as sns
+
+# Load data (ensure the path to your CSV is correct)
+data = pd.read_csv("../data/cleaned_water_quality_data.csv")
+df = pd.DataFrame(data)
+
+# Ensure the collectiondate is in datetime format
+df['collectiondate'] = pd.to_datetime(df['collectiondate'], errors='coerce', format='%Y-%m-%d')
+
+# Add a column for the day of the week
+df['day_of_week'] = df['collectiondate'].dt.day_name()
+
+# Group by day of the week and beachname, then calculate average E. coli levels
+average_ecoli = df.groupby(['day_of_week', 'beachname'])['ecoli'].mean().reset_index()
+
+# Sort days of the week in the typical order
+average_ecoli['day_of_week'] = pd.Categorical(
+    average_ecoli['day_of_week'],
+    categories=['Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday', 'Saturday', 'Sunday'],
+    ordered=True
+)
+
+# Create the bar chart
+plt.figure(figsize=(14, 8))
+sns.barplot(data=average_ecoli, x='day_of_week', y='ecoli', hue='beachname', palette='viridis')
+plt.title('Average E. coli Rates by Day of the Week and Beach', fontsize=16)
+plt.xlabel('Day of the Week', fontsize=14)
+plt.ylabel('Average E. coli Level (MPN/100mL)', fontsize=14)
+plt.xticks(rotation=45)
+plt.grid(axis='y', linestyle='--', linewidth=0.5, alpha=0.7)
+plt.legend(title='Beach Name', bbox_to_anchor=(1.05, 1), loc='upper left')
+plt.tight_layout()
+plt.show()

### 2024-09-20 code/model_generation/dow_scattr.py

@@ -1,0 +1,37 @@
+import pandas as pd
+import matplotlib.pyplot as plt
+import seaborn as sns
+from matplotlib.ticker import MaxNLocator
+
+# Sample data creation (replace this with your actual data)
+data = pd.read_csv("data/cleaned_water_quality_data.csv")
+
+# Create a DataFrame
+df = pd.DataFrame(data)
+
+# Add a column for the day of the week
+df['collectiondate'] = pd.to_datetime(df['collectiondate'])
+df['day_of_week'] = df['collectiondate'].dt.day_name()
+
+# Days of the week order for plotting
+days_of_week = ['Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday', 'Saturday', 'Sunday']
+
+# Create scatter plots for each day of the week
+for day in days_of_week:
+    # Filter data for the specific day of the week
+    day_df = df[df['day_of_week'] == day]
+    
+    # Create the scatter plot
+    plt.figure(figsize=(14, 7))
+    sns.scatterplot(data=day_df, x='collectiondate', y='ecoli', hue='beachname', palette='Set1', s=75, alpha=0.6)
+    plt.title(f'E. coli Levels Over Time - {day}', fontsize=16)
+    plt.xlabel('Collection Date', fontsize=14)
+    plt.ylabel('E. coli Level', fontsize=14)
+    plt.gca().xaxis.set_major_locator(MaxNLocator(nbins=8))
+    plt.gca().xaxis.set_major_formatter(plt.matplotlib.dates.DateFormatter('%b %Y'))
+    plt.xticks(rotation=45, ha='right')
+      # Log scale for better representation
+    plt.grid(visible=True, linestyle='--', linewidth=0.5, alpha=0.7)
+    plt.legend(title='Beach Name', loc='upper left', bbox_to_anchor=(1.05, 1))
+    plt.tight_layout()
+    plt.show()

### 2024-09-20 code/model_generation/ecoli_scatter.py

@@ -1,0 +1,40 @@
+# Import packages
+import pandas as pd
+import tabulate as tab
+import seaborn as sns
+import matplotlib.pyplot as plt
+import numpy as np
+from matplotlib.ticker import MaxNLocator
+
+plt.ion()
+# Import data
+data = pd.read_csv("../data/cleaned_water_quality_data.csv")
+df = pd.DataFrame(data)
+
+# generate graph
+df['collectiondate'] = pd.to_datetime(df['collectiondate'])
+df['log_ecoli'] = np.log(df['ecoli'])
+
+sns.set(style="whitegrid", context="talk")
+
+# Create a scater plot
+sns.scatterplot(data=df, x='collectiondate', y='ecoli', hue='beachname', palette='Set1', s=20, alpha=0.7)
+
+# Set title and labels
+plt.title('E. coli Levels Over Time by Site and Beach', fontsize=16)
+plt.xlabel('Collection Date', fontsize=14)
+plt.ylabel('E. coli Level (MPN/100mL)', fontsize=14)
+plt.grid(True, axis='x')
+# modify x axis labels
+plt.xticks(df['collectiondate'][::100], rotation=45)  
+plt.tight_layout() 
+
+plt.gca().xaxis.set_major_locator(MaxNLocator(nbins=10))
+plt.gca().xaxis.set_major_formatter(plt.matplotlib.dates.DateFormatter('%Y %M'))
+plt.legend(title='Beach Name', loc='upper left', bbox_to_anchor=(1.05, 1))
+
+
+
+plt.savefig("results/ecoli_levels_plot.png", dpi=300, bbox_inches='tight') 
+
+plt.show

### 2024-09-20 code/model_generation/individual_beach_scatter.py

@@ -1,0 +1,42 @@
+
+import pandas as pd
+import matplotlib.pyplot as plt
+import seaborn as sns
+from matplotlib.ticker import MaxNLocator
+
+data = pd.read_csv("../data/cleaned_water_quality_data.csv")
+df = pd.DataFrame(data)
+
+# Separate data by beach
+beach1_df = df[df['beachname'] == 'Sunnyside Beach']
+beach2_df = df[df['beachname'] == 'Marie Curtis Park East Beach']
+
+# Create the plot for Sunnyside Beach
+plt.figure(figsize=(14, 7))
+sns.scatterplot(data=beach1_df, x='collectiondate', y='ecoli', hue='sitename', palette='Set1', s=75, alpha=0.6)
+plt.title('E. coli Levels Over Time - Sunnyside Beach', fontsize=16)
+plt.xlabel('Collection Date', fontsize=14)
+plt.ylabel('E. coli Level', fontsize=14)
+plt.gca().xaxis.set_major_locator(MaxNLocator(nbins=8))
+plt.gca().xaxis.set_major_formatter(plt.matplotlib.dates.DateFormatter('%b %Y'))
+plt.xticks(rotation=45, ha='right')
+  # Log scale for better representation
+plt.grid(visible=True, linestyle='--', linewidth=0.5, alpha=0.7)
+plt.legend(title='Site Name', loc='upper left', bbox_to_anchor=(1.05, 1))
+plt.tight_layout()
+plt.show()
+
+# Create the plot for Marie Curtis Park East Beach
+plt.figure(figsize=(14, 7))
+sns.scatterplot(data=beach2_df, x='collectiondate', y='ecoli', hue='sitename', palette='Set2', s=75, alpha=0.6)
+plt.title('E. coli Levels Over Time - Marie Curtis Park East Beach', fontsize=16)
+plt.xlabel('Collection Date', fontsize=14)
+plt.ylabel('E. coli Level', fontsize=14)
+plt.gca().xaxis.set_major_locator(MaxNLocator(nbins=8))
+plt.gca().xaxis.set_major_formatter(plt.matplotlib.dates.DateFormatter('%b %Y'))
+plt.xticks(rotation=45, ha='right')
+  # Log scale for better representation
+plt.grid(visible=True, linestyle='--', linewidth=0.5, alpha=0.7)
+plt.legend(title='Site Name', loc='upper left', bbox_to_anchor=(1.05, 1))
+plt.tight_layout()
+plt.show()

### 2024-09-20 code/model_generation/summary_tables.py

@@ -66,47 +66,19 @@ def create_ecoli_stats(df):
     ecoli_stats = tab.tabulate(ecoli_stats, headers='keys', tablefmt='github')
     return ecoli_stats
 
-def e_levels_by_site(df):
 
-    df['log_ecoli'] = np.log(df['ecoli'])
-    # Resample by week and calculate the mean ecoli level, grouping by beachname and sitename
-    
-    sns.set(style="whitegrid", context="talk")
-
-    # Create the line plot
-    plt.figure(figsize=(12, 6))
-    sns.lineplot(
-        df,
-        x='collectiondate',
-        y='log_ecoli',
-        hue='beachname',  # Color by beachname
-        style='beachname',  # Style lines by sitename
-        dashes=False,  # Solid lines
-        palette='tab10'  # Optional: Choose a color palette
-    )
-
-    plt.title('E. coli Levels Over Time by Site and Beach')
-    plt.xlabel('Collection Date')
-    plt.ylabel('E. coli Level')
-    plt.xticks(rotation=45)  # Rotate x-axis labels for better readability
-    plt.legend(title='Beach Name', bbox_to_anchor=(1.05, 1), loc='upper left')  # Place legend outside the plot
-    plt.tight_layout() 
-
-    plt.savefig("results/ecoli_levels_plot.png", dpi=300, bbox_inches='tight') 
-    return "graph_1 made"
 
 def main():
     #Import Data
     data = pd.read_csv("data/cleaned_water_quality_data.csv")
     df = pd.DataFrame(data)
 
-    #with open("results/summary_table.md", "w") as file:
-    #    file.write(create_counts(df))
+    with open("results/summary_table.md", "w") as file:
+        file.write(create_counts(df))
     
-    #with open("results/ecoli_stats.md", "w") as file:
-    #    file.write(create_ecoli_stats(df))
+    with open("results/ecoli_stats.md", "w") as file:
+        file.write(create_ecoli_stats(df))
 
-    e_levels_by_site(df)
 
 # Run the script
 if __name__ == "__main__":

### 2024-09-23 code/model_generation/ecoli_scatter.py
converts scatter plot to log

@@ -6,9 +6,8 @@ import matplotlib.pyplot as plt
 import numpy as np
 from matplotlib.ticker import MaxNLocator
 
-plt.ion()
 # Import data
-data = pd.read_csv("../data/cleaned_water_quality_data.csv")
+data = pd.read_csv("data/cleaned_water_quality_data.csv")
 df = pd.DataFrame(data)
 
 # generate graph
@@ -18,19 +17,22 @@ df['log_ecoli'] = np.log(df['ecoli'])
 sns.set(style="whitegrid", context="talk")
 
 # Create a scater plot
-sns.scatterplot(data=df, x='collectiondate', y='ecoli', hue='beachname', palette='Set1', s=20, alpha=0.7)
+sns.scatterplot(data=df, x='collectiondate', y='log_ecoli', hue='beachname', palette='Set1', s=20, alpha=0.7)
 
 # Set title and labels
 plt.title('E. coli Levels Over Time by Site and Beach', fontsize=16)
 plt.xlabel('Collection Date', fontsize=14)
-plt.ylabel('E. coli Level (MPN/100mL)', fontsize=14)
+plt.ylabel('E. coli Level (Log(MPN/100mL))', fontsize=14)
 plt.grid(True, axis='x')
 # modify x axis labels
-plt.xticks(df['collectiondate'][::100], rotation=45)  
-plt.tight_layout() 
+plt.xticks(df['collectiondate'], fontsize=10, rotation=75 )  
+plt.yticks(df['log_ecoli'], fontsize=10)
+plt.tight_layout()
 
-plt.gca().xaxis.set_major_locator(MaxNLocator(nbins=10))
-plt.gca().xaxis.set_major_formatter(plt.matplotlib.dates.DateFormatter('%Y %M'))
+plt.gca().xaxis.set_major_locator(MaxNLocator(nbins=17))
+plt.gca().yaxis.set_major_locator(MaxNLocator(nbins=10))
+
+plt.gca().xaxis.set_major_formatter(plt.matplotlib.dates.DateFormatter('%b %Y'))
 plt.legend(title='Beach Name', loc='upper left', bbox_to_anchor=(1.05, 1))
 
 

### 2024-09-23 code/model_generation/montly_ecoli.py
Creates monthly bar graph

@@ -1,0 +1,31 @@
+import pandas as pd
+import matplotlib.pyplot as plt
+import seaborn as sns
+
+# Load data (ensure the path to your CSV is correct)
+data = pd.read_csv("data/cleaned_water_quality_data.csv")
+df = pd.DataFrame(data)
+
+# Ensure the collectiondate is in datetime format
+df['collectiondate'] = pd.to_datetime(df['collectiondate'], errors='coerce', format='%Y-%m-%d')
+
+# Add a column for the day of the week
+df['month'] = df['collectiondate'].dt.month_name()
+
+# Group by day of the week and beachname, then calculate average E. coli levels
+average_ecoli = df.groupby(['month', 'beachname'])['ecoli'].mean().reset_index()
+
+# Sort days of the week in the typical order
+
+
+# Create the bar chart
+plt.figure(figsize=(14, 8))
+sns.barplot(data=average_ecoli, x='month', y='ecoli', hue='beachname', palette='viridis')
+plt.title('Average E. coli Rates by Day of the Week and Beach', fontsize=16)
+plt.xlabel('Day of the Week', fontsize=14)
+plt.ylabel('Average E. coli Level (MPN/100mL)', fontsize=14)
+plt.xticks(rotation=45)
+plt.grid(axis='y', linestyle='--', linewidth=0.5, alpha=0.7)
+plt.legend(title='Beach Name', bbox_to_anchor=(1.05, 1), loc='upper left')
+plt.tight_layout()
+plt.show()

### 2024-09-23 paper/paper.qmd
Create Paper

@@
+---
+title: "Water Quality at Toronto Beaches"
+subtitle: "An analysis of e-coli rates in toronto beaches 2007-2024"
+author: 
+  - Jacob Gilbert
+thanks: "Code and data are available at: https://github.com/JfpGilbert0/toronto-beaches-water-quality."
+date: today
+date-format: long
+abstract: ""
+format: pdf
+---
+
+```{python}
+#| echo: false
+# import dependancies
+import jupyter
+```
+# Introduction
+
+## Referencing
+
+# Data
+
+## Data Source
+Opendatatoronto is a publicly available collection of wide ranges of data.
+This resource was used to obtain the e-coli levels of the water at two of Toronto's many beaches.
+The data set included coli levels from the SunnySide and Marie Curtis park east beaches, from the 3rd of July 2007 to September 5th 2024.
+The large date range allows us to extrapolate trends well and further the recency of the data allows us to generate a good understanding of the present.
+We have a total of 3084 data points from the two beaches, 1370 and 1714 data points respectively.
+Each detailing the collectiondate, beachname, sitename, geometric location (given by longitude and latitude) and the e coli levels.
+
+(Summary stats)
+
+## Measurement
+E-coli is measured in MLN/100ml a measurement that reflects the ecoli particles in evey 100ml of lake water.
+For our data presentation log values are used to give a better visualisation of the data.
+To obtain an accurate representation of the beach as a whole 12 different locations here used to gather data across the 2 beaches as to remove bias that one testing location may result in.
+the result is that we obtain an unbiased representation of ecoli levels at each beach as a whole over more than a decade.
+Some missing data we encounter however is that although we have the collection date we do not observe the time that data was collected.
+Missing this data may lead to some bias as the time of day could impact the baseline levels of e-coli in the water.
+What this form of measurement does allow us to do is look at the differences in water quality at a date level, a figure() shows there is a good distribution of data from ach day of the week.
+
+# Methods
+Before we look at the daily differences in levels we will look at the data as a whole, observing the distribution of higher and lower levels of ecoli in the water across time as well as by beach.
+Using canada.ca we observe that they define unsafe levels of ecoli in drinking water to ber 88ML/100ml.
+As such we will use this in our analysis to observe the frequency that “unsafe” levels are observed in the beach water.
+Finally we will look at these distributions over the course of a week to answer the question of when water quality is at its best, and as a lake swimmer when is it safest for me to swim at these beaches.
+
+# Results
+```{python}
+#| echo: false
+# Execute the external Python script
+with open("../code/model_generation/ecoli_scatter.py") as script_file:
+    exec(script_file.read())
+```
+
+Figure() displays our data as a whole, we see very widely distributed values of what could be considered high levels of e-coli from 2007 to 2016.
+We see the highest density of data at 10 MLN/100ml but on the higher end levels range from 1000 to 9000.
+Closer to the present we do not see as much variation on the higher end.
+In fact post 2017 the data groups either at the low level of 10 or at around 1000 almost exclusively.
+The variations observed appear to not vary by beach in figure.This is highlighted in the following figures where we see similar distributions of values when separating the data by beach.
+The following two figures alo highlight that the site locations appear in the data fairly randomly supporting that these different sites are not providing much bias.
+```{python}
+#| echo: false
+# Execute the external Python script
+with open("../code/model_generation/individual_beach_scatter.py") as script_file:
+    exec(script_file.read())
+```
+
+## Monthly data
+```{python}
+#| echo: false
+# Execute the external Python script
+#with open(f"../code/model_generation/monthly_summary.py") as script_file:
+#    exec(script_file.read())
+```
+
+Figure 4 looks at the monthly data at each beach, as mentioned in the data section only th summer months provide data due to the cold weather in toronto limiting available data in the winter.
+This data can still provide useful data.
+```{python}
+#| echo: false
+# Execute the external Python script
+with open(f"../code/model_generation/monthly_ecoli.py") as script_file:
+    exec(script_file.read())
+```
+
+## Weekly Data
+
 ```{python}
@@
 
-```{python}
-#| echo: false
-# Execute the external Python script
-with open("../code/model_generation/individual_beach_scatter.py") as script_file:
-    exec(script_file.read())
-```
-
-```{python}
-import pandas as pd
-import matplotlib.pyplot as plt
-import seaborn as sns
-from matplotlib.ticker import MaxNLocator
-
-# Sample data creation (replace this with your actual data)
-data = pd.read_csv("../data/cleaned_water_quality_data.csv")
-
-# Create a DataFrame
-df = pd.DataFrame(data)
-# Add a column for the day of the week
-df['day_of_week'] = df['collectiondate'].dt.day_name()
-
-# Days of the week order for plotting
-days_of_week = ['Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday', 'Saturday', 'Sunday']
-
-# Create scatter plots for each day of the week
-for day in days_of_week:
-    # Filter data for the specific day of the week
-    day_df = df[df['day_of_week'] == day]
-
-    # Create the scatter plot
-    plt.figure(figsize=(14, 7))
-    sns.scatterplot(data=day_df, x='collectiondate', y='ecoli', hue='beachname', palette='Set1', s=75, alpha=0.6)
-    plt.title(f'E. coli Levels Over Time - {day}', fontsize=16)
-    plt.xlabel('Collection Date', fontsize=14)
-    plt.ylabel('E. coli Level', fontsize=14)
-    plt.gca().xaxis.set_major_locator(MaxNLocator(nbins=8))
-    plt.gca().xaxis.set_major_formatter(plt.matplotlib.dates.DateFormatter('%b %Y'))
-    plt.xticks(rotation=45, ha='right')
-      # Log scale for better representation
-    plt.grid(visible=True, linestyle='--', linewidth=0.5, alpha=0.7)
-    plt.legend(title='Beach Name', loc='upper left', bbox_to_anchor=(1.05, 1))
-    plt.tight_layout()
-    plt.show()
-
-```
+# Conclusion

### 2024-09-23 code/model_generation/dow_binary.py
Model creation and alteration for use in Quarto Document

@@ -1,0 +1,50 @@
+import pandas as pd
+import matplotlib.pyplot as plt
+import seaborn as sns
+
+# Load data (ensure the path to your CSV is correct)
+data = pd.read_csv("../data/cleaned_water_quality_data.csv")
+df = pd.DataFrame(data)
+
+# Ensure the collectiondate is in datetime format
+df['collectiondate'] = pd.to_datetime(df['collectiondate'], errors='coerce', format='%Y-%m-%d')
+
+# Add a column for the day of the week
+df['day_of_week'] = df['collectiondate'].dt.day_name()
+
+df['above_88'] = df['ecoli'].apply(lambda x: 1 if x > 88 else 0)
+
+#summary table for dow.
+day_counts = df['day_of_week'].value_counts().reindex([
+    'Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday', 'Saturday', 'Sunday'
+]).reset_index()
+
+# Rename the columns for clarity
+day_counts.columns = ['day_of_week', 'count']
+
+# Display the summary table
+print(day_counts)
+# Group by day of the week and beachname, then calculate average E. coli levels
+average_dangerous = df.groupby(['day_of_week', 'beachname'])['above_88'].mean().reset_index()
+
+# Sort days of the week in the typical order
+average_dangerous['day_of_week'] = pd.Categorical(
+    average_dangerous['day_of_week'],
+    categories=['Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday', 'Saturday', 'Sunday'],
+    ordered=True
+)
+
+# Create binary value for dangeroud e-coli levels
+
+
+# Create the bar chart
+plt.figure(figsize=(14, 8))
+sns.barplot(data=average_dangerous, x='day_of_week', y='above_88', hue='beachname', palette='viridis')
+plt.title('Average times Ecoli levels above 88MLN/100ml', fontsize=15)
+plt.xlabel('Day of the Week', fontsize=14)
+plt.ylabel('Average days, E. coli Level above 88 (MPN/100mL)', fontsize=12)
+plt.xticks(rotation=45)
+plt.grid(axis='y', linestyle='--', linewidth=0.5, alpha=0.7)
+plt.legend(title='Beach Name', bbox_to_anchor=(1.05, 1), loc='upper left')
+plt.tight_layout()
+plt.show()

### 2024-09-23 code/model_generation/ecoli_scatter.py

@@ -7,7 +7,7 @@ import numpy as np
 from matplotlib.ticker import MaxNLocator
 
 # Import data
-data = pd.read_csv("data/cleaned_water_quality_data.csv")
+data = pd.read_csv("../data/cleaned_water_quality_data.csv")
 df = pd.DataFrame(data)
 
 # generate graph
@@ -34,9 +34,6 @@ plt.gca().yaxis.set_major_locator(MaxNLocator(nbins=10))
 
 plt.gca().xaxis.set_major_formatter(plt.matplotlib.dates.DateFormatter('%b %Y'))
 plt.legend(title='Beach Name', loc='upper left', bbox_to_anchor=(1.05, 1))
-
-
-
-plt.savefig("results/ecoli_levels_plot.png", dpi=300, bbox_inches='tight') 
+ 
 
 plt.show

### 2024-09-23 code/model_generation/individual_beach_scatter.py

@@ -7,6 +7,7 @@ from matplotlib.ticker import MaxNLocator
 data = pd.read_csv("../data/cleaned_water_quality_data.csv")
 df = pd.DataFrame(data)
 
+df['collectiondate'] = pd.to_datetime(df['collectiondate'])
 # Separate data by beach
 beach1_df = df[df['beachname'] == 'Sunnyside Beach']
 beach2_df = df[df['beachname'] == 'Marie Curtis Park East Beach']

### 2024-09-23 code/model_generation/monthly_ecoli.py

@@ -3,7 +3,7 @@ import matplotlib.pyplot as plt
 import seaborn as sns
 
 # Load data (ensure the path to your CSV is correct)
-data = pd.read_csv("data/cleaned_water_quality_data.csv")
+data = pd.read_csv("../data/cleaned_water_quality_data.csv")
 df = pd.DataFrame(data)
 
 # Ensure the collectiondate is in datetime format

### 2024-09-23 code/model_generation/summary_tables.py

@@ -6,10 +6,10 @@ import matplotlib.pyplot as plt
 import numpy as np
 
 # Import data
-data = pd.read_csv("data/cleaned_water_quality_data.csv")
+data = pd.read_csv("../data/cleaned_water_quality_data.csv")
 df = pd.DataFrame(data)
 
-
+df['collectiondate'] = pd.to_datetime(df['collectiondate'])
 # Summary stats tables
 def create_counts(data):
     df = data
@@ -78,6 +78,8 @@ def main():
     
     with open("results/ecoli_stats.md", "w") as file:
         file.write(create_ecoli_stats(df))
+
+
 
 
 # Run the script

### 2024-09-23 paper/paper.qmd
Creae and add to quarto weekly and monthly summary tables

@@
 format: pdf
+jupyter: python3
 ---
@@
 # Execute the external Python script
-#with open(f"../code/model_generation/monthly_summary.py") as script_file:
-#    exec(script_file.read())
+with open(f"../code/model_generation/monthly_summary.py") as script_file:
+    exec(script_file.read())
+summary
 ```
@@
 ## Weekly Data
-
+```{python}
+#| echo: false
+# Execute the external Python script
+with open(f"../code/model_generation/weekly_summary.py") as script_file:
+    exec(script_file.read())
+summary
+```
 ```{python}

### 2024-09-23 code/model_generation/monthly_summary.py

@@ -1,0 +1,24 @@
+
+# Imports
+import pandas as pd
+
+# Load data (ensure the path to your CSV is correct)
+data = pd.read_csv("../data/cleaned_water_quality_data.csv")
+df = pd.DataFrame(data)
+
+# Ensure the collectiondate is in datetime format
+df['collectiondate'] = pd.to_datetime(df['collectiondate'], errors='coerce', format='%Y-%m-%d')
+
+# Extract month and year for grouping
+df['month'] = df['collectiondate'].dt.month_name()  # Extracts month in 'YYYY-MM' format
+
+# Group by month and calculate required statistics
+summary = df.groupby('month').agg(
+    number_of_datapoints=('ecoli', 'size'),
+    average_ecoli=('ecoli', 'mean'),
+    variance=('ecoli', 'var'),
+    number_of_points_above_100=('ecoli', lambda x: (x > 100).sum())
+).reset_index()
+
+# Display the summary table
+summary

### 2024-09-23 code/model_generation/weekly_summary.py

@@ -1,0 +1,19 @@
+import pandas as pd
+
+# Load data (ensure the path to your CSV is correct)
+data = pd.read_csv("../data/cleaned_water_quality_data.csv")
+df = pd.DataFrame(data)
+
+# Ensure the collectiondate is in datetime format
+df['collectiondate'] = pd.to_datetime(df['collectiondate'], errors='coerce', format='%Y-%m-%d')
+
+# Extract month and year for grouping
+df['day'] = df['collectiondate'].dt.day_name()  # Extracts month in 'YYYY-MM' format
+
+# Group by month and calculate required statistics
+summary = df.groupby('day').agg(
+    number_of_datapoints=('ecoli', 'size'),
+    average_ecoli=('ecoli', 'mean'),
+    variance=('ecoli', 'var'),
+    number_of_points_above_100=('ecoli', lambda x: (x > 100).sum())
+).reset_index()

### 2024-09-24 paper/paper.qmd
Add inroduction to qmd file

@@
 ```
+
 # Introduction
+Water contamination is important in Toronto, especially in the summer months as locals flock to the nearby beaches as the weather turns from snow to sun.
+E-coli testing is common at most public swimming locations as high levels highlight risk to those embarking into the water.
+While E-coli is unlikely to cause illness on it's own, If a water sample has high concentrations of E.
+coli, other more dangerous and infectious organisms may be present.(https://www.cleanlakesalliance.org/e-coli/).
+Thus the levels at public beaches can affect public health, and as such action should be taken on the results.
+To either close public beaches that exhibit unsafe levels, or continuation of previous action that has led to lower levels.
+Despite these public health measures, there is limited understanding of the specific patterns of E.
+coli contamination in Toronto's beaches.
+Most public advisories and beach management decisions rely on reactive measures rather than proactive analysis of historical trends.
+This paper aims to fill this gap by examining historical E.
+coli data from Open Data Toronto, focusing on two of the city’s popular beaches: Sunnyside Beach and Marie Curtis Park East Beach.
+By analyzing data over several years, this study seeks to identify temporal patterns in contamination, such as variations by day of the week and season, to help answer the critical question: When are beachgoers at the highest risk of swimming in contaminated waters?
+Additionally, this analysis explores whether E.
+coli levels have shown improvement over time, informing on the effectiveness of current water management practices.
+Understanding the current and historic trends are paramount for city officials to understand if current water quality measures are sufficient.
+Or in the worse case, that interventions are needed in order to keep Toronto beaches safe for everyone.
+This paper is structured as follows: First, we present the dataset, describing its sources, key variables, and the data cleaning process.
+Next, we analyse patterns in E.
+coli levels, examining how contamination varies across different times and locations.
+We then discuss the implications of these findings for public health and beach management, followed by a conclusion that summarises the key insights and recommendations for future action.
 
 ## Referencing
+In this paper, Python [] and it packages are utilized as the primary programming language for the following analysis and displaying the data.
+Key packages used include pandas [McKinney, 2010], which facilitated efficient data cleaning, manipulation, and summarization of the dataset, NumPy [Harris et al., 2020] was employed for numerical operations, Matplotlib [Hunter, 2007] and seaborn [Waskom, 2021] were critical for creating informative visualizations.
 

### 2024-09-25 paper/paper.qmd
First draft of results section

@@
 This data can still provide useful data.
+May exhibits the lowest average E.
+coli levels, at 101.76 MPN/100mL, which is just above the unsafe threshold.
+With only 6.16% of samples being unsafe.
+The low variance suggests that water quality is fairly consistent during this early summer month.
+August and July show extremely high avrages, above 500MPN/100ml, with over 30% of daa collected here showing water quality above acceptable levels.
 ```{python}
@@
 ```
+Figure X summarises E.
+coli data by day of the week, including the total number of samples (n), average E.
+coli levels (average_ecoli), variance, the number of samples with unsafe levels (unsafe_levels), and the percentage of samples deemed unsafe (percentage_unsafe).
+The weekdays, Monday through Friday, see the highest average amounts of e-coli, as well as the higher probability that the water is deemed unsafe.
+Tuesday and Wednesday see average levels over 5x the 100 MPN/100ml unsafe , with relatively low variance highlighting these days as the worst for water quality over the course of a week.
+This point is further supported with both days having around 30% of their levels above the threshold.
+The weekend encounters the lowest levels by comparison, 134.91 MPN/100mL and 196.34 MPN/100mL on saturday and sunday respectively.
+with under 10% of tests having an unsafe result.
+This Is a positive sight as these days, it is assumed, encounter the most activity due to the 5 day work week.
+Figure() gives context to how the two beaches differ in their levels over the course of a week.
+Marie Curtis see's fairly consistent averages across the week, between 300 and 600 MPN/100ml.
+We still observe the lower average observed in the table here however it is far more significant at this location on Saturday compared to Sunday.
+As is clear from the graph we see significant differences in the average in the two locations.
+Sunnyside Beach shows extraordinarily high levels on Monday and Tuesday Compared to Marie Curtis.
+The rest of the weekdays we observe a fairly insignificant difference between the locations average level, but Sunnyside does appear to have a better average quality.
+The weekend we see the inverse to the start of the week, with Sunnyside having very low averages.
+In fact Sunday at Sunnyside is the only category here that has average levels below the safe threshold of 100 MPN/100ml.
+
 ```{python}

### 2024-09-25 code/model_generation/monthly_summary.py

@@ -14,11 +14,13 @@ df['month'] = df['collectiondate'].dt.month_name()  # Extracts month in 'YYYY-MM
 
 # Group by month and calculate required statistics
 summary = df.groupby('month').agg(
-    number_of_datapoints=('ecoli', 'size'),
+    n=('ecoli', 'size'),
     average_ecoli=('ecoli', 'mean'),
     variance=('ecoli', 'var'),
-    number_of_points_above_100=('ecoli', lambda x: (x > 100).sum())
+    unsafe_levels=('ecoli', lambda x: (x > 100).sum())
 ).reset_index()
-
-# Display the summary table
-summary
+summary['percentage_unsafe'] = round((summary['unsafe_levels']/summary['n'])*100, 2)
+# order by month
+months = ['May', 'June', 'July', 'August', 'September']
+summary['month'] = pd.Categorical(summary['month'], categories=months, ordered=True)
+summary = summary.sort_values(by='month').reset_index(drop=True)

### 2024-09-25 code/model_generation/summary_tables.py

@@ -73,10 +73,10 @@ def main():
     data = pd.read_csv("data/cleaned_water_quality_data.csv")
     df = pd.DataFrame(data)
 
-    with open("results/summary_table.md", "w") as file:
+    with open("results/tables/summary_table.md", "w") as file:
         file.write(create_counts(df))
     
-    with open("results/ecoli_stats.md", "w") as file:
+    with open("results/tables/ecoli_stats.md", "w") as file:
         file.write(create_ecoli_stats(df))
 
 

### 2024-09-25 code/model_generation/weekly_summary.py

@@ -12,8 +12,15 @@ df['day'] = df['collectiondate'].dt.day_name()  # Extracts month in 'YYYY-MM' fo
 
 # Group by month and calculate required statistics
 summary = df.groupby('day').agg(
-    number_of_datapoints=('ecoli', 'size'),
+    n=('ecoli', 'size'),
     average_ecoli=('ecoli', 'mean'),
     variance=('ecoli', 'var'),
-    number_of_points_above_100=('ecoli', lambda x: (x > 100).sum())
+    unsafe_levels=('ecoli', lambda x: (x > 100).sum())
 ).reset_index()
+summary.index = range(1,8)
+summary['percentage_unsafe'] = round((summary['unsafe_levels']/summary['n'])*100, 2)
+# order by dow
+days = ['Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday', 'Saturday', 'Sunday']
+summary['day'] = pd.Categorical(summary['day'], categories=days, ordered=True)
+summary = summary.sort_values(by='day').reset_index(drop=True)
+print(summary)

### 2024-09-25 code/model_generation/yearly_percentages.py
create yearly percentages graph

@@ -1,0 +1,36 @@
+# Imports
+from matplotlib import pyplot as plt
+import pandas as pd
+import seaborn as sns
+
+# Load data (ensure the path to your CSV is correct)
+data = pd.read_csv("../data/cleaned_water_quality_data.csv")
+df = pd.DataFrame(data)
+
+# Ensure the collectiondate is in datetime format
+df['collectiondate'] = pd.to_datetime(df['collectiondate'], errors='coerce', format='%Y-%m-%d')
+
+# Extract year for grouping
+df['year'] = df['collectiondate'].dt.year  # Extracts month in 'YYYY-MM' format
+
+# Group by year and calculate required statistics
+yearly_avg = df.groupby('year').agg(
+    n=('ecoli', 'size'),
+    average_ecoli=('ecoli', 'mean'),
+    variance=('ecoli', 'var'),
+    unsafe_levels=('ecoli', lambda x: (x > 100).sum())
+).reset_index()
+yearly_avg['percentage_unsafe'] = round((yearly_avg['unsafe_levels']/yearly_avg['n'])*100, 2)
+yearly_avg = yearly_avg.sort_values(by='year').reset_index(drop=True)
+# Create Graph
+# Create the scatter plot
+plt.figure(figsize=(14, 7))
+sns.barplot(data=yearly_avg, x='year', y='percentage_unsafe', alpha=0.6)
+plt.title(f'Percentage of tests above 100MPN/100ml By Year', fontsize=16)
+plt.xlabel('Year', fontsize=14)
+plt.ylabel('Percentage of unsafe ecoli levels', fontsize=10)
+plt.xticks(rotation=45, ha='right')
+plt.tight_layout()
+
+plt.show()
+

### 2024-09-25 paper/paper.qmd
paper second draft

@@
 Missing this data may lead to some bias as the time of day could impact the baseline levels of e-coli in the water.
-What this form of measurement does allow us to do is look at the differences in water quality at a date level, a figure() shows there is a good distribution of data from ach day of the week.
+What this form of measurement does allow us to do is look at the differences in water quality at a date level, a figure() shows there is a good distribution of data from each day of the week.
+Ther are some gaps in the data however, in ehe monthly data we only find rsults from the summer months.
+As during the summer the impact of water quality is at its highest this doesn't impact the usefulnss of the data, howevr it does result in missing context of levels during the rest of the year.
 
@@
 # Results
+## Time Series
 ```{python}
@@
 In fact post 2017 the data groups either at the low level of 10 or at around 1000 almost exclusively.
-The variations observed appear to not vary by beach in figure.This is highlighted in the following figures where we see similar distributions of values when separating the data by beach.
-The following two figures alo highlight that the site locations appear in the data fairly randomly supporting that these different sites are not providing much bias.
+The variations observed appear to not vary by beach in figure.
+
+Figure 1 is a scatter plot of E.
+coli levels over time by site and beach, and displays the data as a whole.
+It highlights the trends in contamination levels at Sunnyside Beach and Marie Curtis Park East Beach from 2008 to 2024.the beach site itself does not seem to significantly impact the overall trend of E.
+coli levels over time, as both Sunnyside and Marie Curtis Park East show similar patterns of fluctuation.
+the magnitude of the fluctuations ar very notable as prior to 2017 there are spikes in e-coli levels to many degrees, some extremely high.
+Whereas afer 2017 we see levels almost exclusivly sitting at 10 or around 1000 MNP/100ml.
+This appears to indicate that although the risks are persistant over time they are lessend in recent years.
+
+Figure 2 is a bar chart illustrating the percentage of tests above the unsafe threshold by year, provides an important contrast to the scatter plot.
+Although the scatter plot suggests a reduction in extreme E.
+coli levels in the later years, this bar graph reveals that the frequency of tests exceeding the 100 MPN/100mL threshold does not diminish in the same way.
+Years such as 2010, 2011, 2019, and 2021 see substantial spikes in the percentage of unsafe tests, reaching as high as over 50% in 2010.
+We do not see percentages moving consitently in on direction other than after 2017 where the rates appear to be climbing.
+2017 is a strang anomoly here with no tests returning unsafe results.
+The combination of these two graphs paint a clear picture of water quality in Toronto over the years.
+
 ```{python}
 #| echo: false
-# Execute the external Python script
-with open("../code/model_generation/individual_beach_scatter.py") as script_file:
+#| label: fig-2
+# Execute the external Python script for individual beach scatter plots
+with open("../code/model_generation/yearly_percentages.py") as script_file:
     exec(script_file.read())
 ```
+
+**Figure 2**: Bar graph reflecting the percentage of results each year where the water quality exceeds 100MPN/100ml.
+There is no obvious pattern here although in later years we see a trend upwards and rates appear to trend down until 2017 where the rate is 0.
 
@@
 #| echo: false
-# Execute the external Python script
-with open(f"../code/model_generation/monthly_summary.py") as script_file:
+#| label: fig-3
+# Execute the external Python script for monthly summary statistics
+with open("../code/model_generation/monthly_summary.py") as script_file:
     exec(script_file.read())
@@
 
+**Figure 3**: Monthly summary statistics indicate May exhibits the lowest average E. coli levels, while July and August have the highest, with over 30% of samples deemed unsafe.
+
 Figure 4 looks at the monthly data at each beach, as mentioned in the data section only th summer months provide data due to the cold weather in toronto limiting available data in the winter.
-This data can still provide useful data.
+Further we see a disproportionate number of daa points in May and September, this could make the results here less reliable.
+However this data can still provide useful data.
 May exhibits the lowest average E.
@@
 The low variance suggests that water quality is fairly consistent during this early summer month.
-August and July show extremely high avrages, above 500MPN/100ml, with over 30% of daa collected here showing water quality above acceptable levels.
+August and July show extremely high averages, above 500MPN/100ml, with over 30% of daa collected here showing water quality above acceptable levels.
+Looking at figure() these high levels are consistent between both beaches, though we do observe extremely high average levels at Sunnyside.
+Overall, we see the average level aswell a the likelyhood of unsafe levels increase as the summer progresses and slighly taper off in september.
+Although the absence of a reasonable amount of data in September could contribute this increasing trend.
+
 ```{python}
@@
 #| echo: false
-# Execute the external Python script
-with open(f"../code/model_generation/weekly_summary.py") as script_file:
+#| label: fig-5
+# Execute the external Python script for weekly summary statistics
+with open("../code/model_generation/weekly_summary.py") as script_file:
     exec(script_file.read())
@@
 ```
+
+**Figure 5**: Weekly summary statistics show the highest E. coli levels and unsafe probabilities on weekdays, particularly Tuesday and Wednesday, with much lower levels observed on weekends.
+
 Figure X summarises E.
@@
 #| echo: false
-# Execute the external Python script
-with open(f"../code/model_generation/dow_bar.py") as script_file:
+#| label: fig-6
+# Execute the external Python script for weekly E. coli levels by beach
+with open("../code/model_generation/dow_bar.py") as script_file:
     exec(script_file.read())
@@
 
+**Figure 6**: E.
+coli levels by day of the week show significant differences between the two beaches, with Sunnyside exhibiting extraordinarily high levels on Monday and Tuesday compared to Marie Curtis.
+
 # Conclusion
+This analysis of E.
+coli contamination levels at Sunnyside Beach and Marie Curtis Park East Beach reveals critical insights into the patterns and trends of water quality in Toronto’s public swimming areas.
+While the data shows a reduction in the variability of E.
+coli levels in recent years, the frequency of unsafe readings remains concerning, particularly during peak summer months.
+The persistent risk highlighted by the bar graph indicates that ongoing contamination remains a public health challenge, despite efforts to improve water quality.
+Policy makers should improve their effors to reduce the number of days that water quality is unacceptable over the summer months.
+The months that are exhibiting the greates risk are..

### 2024-09-26 paper/paper.qmd
Paper draft 3 (adds references also)

@@
 date-format: long
-abstract: ""
+abstract: "This study analyses E.
+coli contamination trends at two of Toronto's popular beaches, Sunnyside Beach and Marie Curtis Park East Beach, from 2007 to 2024.
+Using data sourced from Open Data Toronto, the analysis examines temporal patterns in contamination levels, assessing the frequency and severity of unsafe E.
+coli concentrations as defined by public health thresholds.
+The study reveals that while E.
+coli levels appear less erratic and generally lower in magnitude in recent years, the frequency of tests exceeding unsafe thresholds has not significantly diminished, indicating persistent contamination risks.
+These findings underscore the need for continued monitoring and targeted interventions to protect public health.
+This research aims to inform beach management practices and guide future efforts to improve water quality at Toronto’s beaches, Aswell as provide insights for swimmers on the risks that they are exposed to."
 format: pdf
 jupyter: python3
+bibliography: references.bib
 ---
@@
 #| echo: false
-# import dependancies
+# import dependencies
 import jupyter
@@
 E-coli testing is common at most public swimming locations as high levels highlight risk to those embarking into the water.
-While E-coli is unlikely to cause illness on it's own, If a water sample has high concentrations of E.
-coli, other more dangerous and infectious organisms may be present.(https://www.cleanlakesalliance.org/e-coli/).
+While E-coli is unlikely to cause illness on its own, If a water sample has high concentrations of E.
+coli, other more dangerous and infectious organisms may be present [@cleanlakesalliance].
 Thus the levels at public beaches can affect public health, and as such action should be taken on the results.
@@
 coli data from Open Data Toronto, focusing on two of the city’s popular beaches: Sunnyside Beach and Marie Curtis Park East Beach.
-By analyzing data over several years, this study seeks to identify temporal patterns in contamination, such as variations by day of the week and season, to help answer the critical question: When are beachgoers at the highest risk of swimming in contaminated waters?
+By analysing data over several years, this study seeks to identify temporal patterns in contamination, such as variations by day of the week and season, to help answer the critical question: When are beachgoers at the highest risk of swimming in contaminated waters?
 Additionally, this analysis explores whether E.
@@
 ## Referencing
-In this paper, Python [] and it packages are utilized as the primary programming language for the following analysis and displaying the data.
-Key packages used include pandas [McKinney, 2010], which facilitated efficient data cleaning, manipulation, and summarization of the dataset, NumPy [Harris et al., 2020] was employed for numerical operations, Matplotlib [Hunter, 2007] and seaborn [Waskom, 2021] were critical for creating informative visualizations.
+In this paper, Python [@python] and its packages are utilised as the primary programming language for the following analysis and displaying the data.
+Key packages used include pandas [McKinney, 2010], which facilitated efficient data cleaning, manipulation, and summarization of the dataset, NumPy [@numpy] was employed for numerical operations, Matplotlib [matplotlib] and seaborn [@seaborn] were critical for creating informative visualisations.
 
@@
 The data set included coli levels from the SunnySide and Marie Curtis park east beaches, from the 3rd of July 2007 to September 5th 2024.
-The large date range allows us to extrapolate trends well and further the recency of the data allows us to generate a good understanding of the present.
+The large data range allows us to extrapolate trends well and further the recency of the data allows us to generate a good understanding of the present.
 We have a total of 3084 data points from the two beaches, 1370 and 1714 data points respectively.
@@
 ## Measurement
-E-coli is measured in MLN/100ml a measurement that reflects the ecoli particles in evey 100ml of lake water.
+E-coli is measured in MPN/100ml a measurement that reflects the e coli particles in every 100ml of lake water.
 For our data presentation log values are used to give a better visualisation of the data.
 To obtain an accurate representation of the beach as a whole 12 different locations here used to gather data across the 2 beaches as to remove bias that one testing location may result in.
-the result is that we obtain an unbiased representation of ecoli levels at each beach as a whole over more than a decade.
+The result is that we obtain an unbiased representation of e coli levels at each beach as a whole over more than a decade.
 Some missing data we encounter however is that although we have the collection date we do not observe the time that data was collected.
@@
 What this form of measurement does allow us to do is look at the differences in water quality at a date level, a figure() shows there is a good distribution of data from each day of the week.
-Ther are some gaps in the data however, in ehe monthly data we only find rsults from the summer months.
-As during the summer the impact of water quality is at its highest this doesn't impact the usefulnss of the data, howevr it does result in missing context of levels during the rest of the year.
+There are some gaps in the data however, in the monthly data we only find results from the summer months.
+As during the summer the impact of water quality is at its highest this doesn't impact the usefulness of the data, however it does result in missing context of levels during the rest of the year.
 
 # Methods
-Before we look at the daily differences in levels we will look at the data as a whole, observing the distribution of higher and lower levels of ecoli in the water across time as well as by beach.
-Using canada.ca we observe that they define unsafe levels of ecoli in drinking water to ber 88ML/100ml.
-As such we will use this in our analysis to observe the frequency that “unsafe” levels are observed in the beach water.
+Before we look at the daily differences in levels we will look at the data as a whole, observing the distribution of higher and lower levels of e coli in the water across time as well as by beach.
+The Ontario government sets their standard of 100 mpn/100ml as their water quality objectives [@ontariowater].
+Thus we will use this as our benchmark for safe e-coli levels in the Beach water.
 Finally we will look at these distributions over the course of a week to answer the question of when water quality is at its best, and as a lake swimmer when is it safest for me to swim at these beaches.
@@
 #| echo: false
+#| label: fig-1
 # Execute the external Python script
@@
 ```
+**Figure 1** depicts E.
+coli contamination levels over time at Sunnyside Beach and Marie Curtis Park East Beach from 2008 to 2024.
+The scatter plot shows that E.
+coli levels were highly variable and frequently above the unsafe threshold in the early years, especially between 2008 and 2013.
+From 2017 onwards, E.
+coli levels appear less erratic, with fewer instances of extreme spikes
 
@@
 coli levels over time, as both Sunnyside and Marie Curtis Park East show similar patterns of fluctuation.
-the magnitude of the fluctuations ar very notable as prior to 2017 there are spikes in e-coli levels to many degrees, some extremely high.
-Whereas afer 2017 we see levels almost exclusivly sitting at 10 or around 1000 MNP/100ml.
-This appears to indicate that although the risks are persistant over time they are lessend in recent years.
+The magnitude of the fluctuations are very notable as prior to 2017 there are spikes in e-coli levels to many degrees, some extremely high.
+Whereas after 2017 we see levels almost exclusively sitting at 10 or around 1000 MNP/100ml.
+This appears to indicate that although the risks are persistent over time they are lessened in recent years.
 
-Figure 2 is a bar chart illustrating the percentage of tests above the unsafe threshold by year, provides an important contrast to the scatter plot.
+Figure 2 is a bar chart illustrating the percentage of tests above the unsafe threshold by year, providing an important contrast to the scatter plot.
 Although the scatter plot suggests a reduction in extreme E.
@@
 Years such as 2010, 2011, 2019, and 2021 see substantial spikes in the percentage of unsafe tests, reaching as high as over 50% in 2010.
-We do not see percentages moving consitently in on direction other than after 2017 where the rates appear to be climbing.
-2017 is a strang anomoly here with no tests returning unsafe results.
+We do not see percentages moving consistently in one direction other than after 2017 where the rates appear to be climbing.
+2017 is a strange anomaly here with no tests returning unsafe results.
 The combination of these two graphs paint a clear picture of water quality in Toronto over the years.
@@
 #| label: fig-2
-# Execute the external Python script for individual beach scatter plots
 with open("../code/model_generation/yearly_percentages.py") as script_file:
@@
 
-**Figure 2**: Bar graph reflecting the percentage of results each year where the water quality exceeds 100MPN/100ml.
+**Figure 2**: Bar graph reflecting the percentage of results each year where the water quality exceeds 100 MPN/100ml.
 There is no obvious pattern here although in later years we see a trend upwards and rates appear to trend down until 2017 where the rate is 0.
@@
 #| echo: false
+#| fig.pos: "H"
 #| label: fig-3
@@
 
-Figure 4 looks at the monthly data at each beach, as mentioned in the data section only th summer months provide data due to the cold weather in toronto limiting available data in the winter.
-Further we see a disproportionate number of daa points in May and September, this could make the results here less reliable.
+Figure 4 looks at the monthly data at each beach, as mentioned in the data section only the summer months provide data due to the cold weather in Toronto limiting available data in the winter.
+Further we see a disproportionate number of data points in May and September, this could make the results here less reliable.
 However this data can still provide useful data.
@@
 The low variance suggests that water quality is fairly consistent during this early summer month.
-August and July show extremely high averages, above 500MPN/100ml, with over 30% of daa collected here showing water quality above acceptable levels.
+August and July show extremely high averages, above 500 MPN/100ml, with over 30% of data collected here showing water quality above acceptable levels.
 Looking at figure() these high levels are consistent between both beaches, though we do observe extremely high average levels at Sunnyside.
-Overall, we see the average level aswell a the likelyhood of unsafe levels increase as the summer progresses and slighly taper off in september.
+Overall, we see the average level as well as the likelihood of unsafe levels increase as the summer progresses and slightly taper off in september.
 Although the absence of a reasonable amount of data in September could contribute this increasing trend.
@@
 #| echo: false
+#| label: fig-4
+#| fig.pos: "H"
 # Execute the external Python script
@@
 
+**figure 4** A bar graph showing the percent of results that are above the government target of 100 mpn/100ml grouped by month.
+The rates peak at both beaches in the middle months.
+Sunnysde experienced the highest rates of the two locations but also the lowest.
+Marie curtis had a bell-curve type of trend over the 4 summer months peaking in July.
+
+\newpage
 ## Weekly Data
@@
 #| echo: false
+#| fig.pos: "H"
 #| label: fig-5
@@
 
+\newpage
 ```{python}
 #| echo: false
+#| fig.pos: "H"
 #| label: fig-6
@@
 ```
-
 **Figure 6**: E.
@@
 The persistent risk highlighted by the bar graph indicates that ongoing contamination remains a public health challenge, despite efforts to improve water quality.
-Policy makers should improve their effors to reduce the number of days that water quality is unacceptable over the summer months.
-The months that are exhibiting the greates risk are..
+Policy makers should improve their efforts to reduce the number of days that water quality is unacceptable over the summer months.
+The months that are exhibiting the greatest risk are July and August, as such Policy makers should focus their efforts on this time of the year.
+As these months fall in the middle of the summer the higher levels could be a result of climate or increased usage of the beaches in the warmer weather.
+From the perspective of swimmers the results displayed here give insights on how to lower the risks of visiting the beaches when water quality is lower.
+Firstly overall we see that in recent years the extremes are much lower which is good news for lake swimmers in toronto.
+We also see that choosing to swim in the earlier months of the summer appear to bring with them better overall water quality than later months.
+Similarly the weekends bear a far lower risk to swimmers, which is fortunate as it is likely that these are the beaches busier times.
+The weekly data does have warning signs also, avoiding Sunnyside during the week, interestingly although this beach has the highest risk days it also pairs that with the safest days in the week.
+
+\newpage

### 2024-09-27 code/model_generation/dow_bar.py
Update py files and include saving functionality to figures and tables

@@ -11,7 +11,7 @@ df['collectiondate'] = pd.to_datetime(df['collectiondate'], errors='coerce', for
 
 # Add a column for the day of the week
 df['day_of_week'] = df['collectiondate'].dt.day_name()
-
+df = df.loc[df['collectiondate'] > '2017-12-31']
 # Group by day of the week and beachname, then calculate average E. coli levels
 average_ecoli = df.groupby(['day_of_week', 'beachname'])['ecoli'].mean().reset_index()
 
@@ -32,4 +32,6 @@ plt.xticks(rotation=45)
 plt.grid(axis='y', linestyle='--', linewidth=0.5, alpha=0.7)
 plt.legend(title='Beach Name', bbox_to_anchor=(1.05, 1), loc='upper left')
 plt.tight_layout()
+
+plt.savefig('../results/figures/ecoli_day_of_week.png', dpi=300, bbox_inches='tight')
 plt.show()

### 2024-09-27 code/model_generation/dow_binary.py

@@ -1,50 +1,0 @@
-import pandas as pd
-import matplotlib.pyplot as plt
-import seaborn as sns
-
-# Load data (ensure the path to your CSV is correct)
-data = pd.read_csv("../data/cleaned_water_quality_data.csv")
-df = pd.DataFrame(data)
-
-# Ensure the collectiondate is in datetime format
-df['collectiondate'] = pd.to_datetime(df['collectiondate'], errors='coerce', format='%Y-%m-%d')
-
-# Add a column for the day of the week
-df['day_of_week'] = df['collectiondate'].dt.day_name()
-
-df['above_88'] = df['ecoli'].apply(lambda x: 1 if x > 88 else 0)
-
-#summary table for dow.
-day_counts = df['day_of_week'].value_counts().reindex([
-    'Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday', 'Saturday', 'Sunday'
-]).reset_index()
-
-# Rename the columns for clarity
-day_counts.columns = ['day_of_week', 'count']
-
-# Display the summary table
-print(day_counts)
-# Group by day of the week and beachname, then calculate average E. coli levels
-average_dangerous = df.groupby(['day_of_week', 'beachname'])['above_88'].mean().reset_index()
-
-# Sort days of the week in the typical order
-average_dangerous['day_of_week'] = pd.Categorical(
-    average_dangerous['day_of_week'],
-    categories=['Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday', 'Saturday', 'Sunday'],
-    ordered=True
-)
-
-# Create binary value for dangeroud e-coli levels
-
-
-# Create the bar chart
-plt.figure(figsize=(14, 8))
-sns.barplot(data=average_dangerous, x='day_of_week', y='above_88', hue='beachname', palette='viridis')
-plt.title('Average times Ecoli levels above 88MLN/100ml', fontsize=15)
-plt.xlabel('Day of the Week', fontsize=14)
-plt.ylabel('Average days, E. coli Level above 88 (MPN/100mL)', fontsize=12)
-plt.xticks(rotation=45)
-plt.grid(axis='y', linestyle='--', linewidth=0.5, alpha=0.7)
-plt.legend(title='Beach Name', bbox_to_anchor=(1.05, 1), loc='upper left')
-plt.tight_layout()
-plt.show()

### 2024-09-27 code/model_generation/dow_scattr.py

@@ -1,37 +1,0 @@
-import pandas as pd
-import matplotlib.pyplot as plt
-import seaborn as sns
-from matplotlib.ticker import MaxNLocator
-
-# Sample data creation (replace this with your actual data)
-data = pd.read_csv("data/cleaned_water_quality_data.csv")
-
-# Create a DataFrame
-df = pd.DataFrame(data)
-
-# Add a column for the day of the week
-df['collectiondate'] = pd.to_datetime(df['collectiondate'])
-df['day_of_week'] = df['collectiondate'].dt.day_name()
-
-# Days of the week order for plotting
-days_of_week = ['Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday', 'Saturday', 'Sunday']
-
-# Create scatter plots for each day of the week
-for day in days_of_week:
-    # Filter data for the specific day of the week
-    day_df = df[df['day_of_week'] == day]
-    
-    # Create the scatter plot
-    plt.figure(figsize=(14, 7))
-    sns.scatterplot(data=day_df, x='collectiondate', y='ecoli', hue='beachname', palette='Set1', s=75, alpha=0.6)
-    plt.title(f'E. coli Levels Over Time - {day}', fontsize=16)
-    plt.xlabel('Collection Date', fontsize=14)
-    plt.ylabel('E. coli Level', fontsize=14)
-    plt.gca().xaxis.set_major_locator(MaxNLocator(nbins=8))
-    plt.gca().xaxis.set_major_formatter(plt.matplotlib.dates.DateFormatter('%b %Y'))
-    plt.xticks(rotation=45, ha='right')
-      # Log scale for better representation
-    plt.grid(visible=True, linestyle='--', linewidth=0.5, alpha=0.7)
-    plt.legend(title='Beach Name', loc='upper left', bbox_to_anchor=(1.05, 1))
-    plt.tight_layout()
-    plt.show()

### 2024-09-27 code/model_generation/dow_summary.py

@@ -19,8 +19,12 @@ summary = df.groupby('day').agg(
 ).reset_index()
 summary.index = range(1,8)
 summary['percentage_unsafe'] = round((summary['unsafe_levels']/summary['n'])*100, 2)
+summary['variance'] = round(summary['variance'], 0)
+summary['average_ecoli']=round(summary['average_ecoli'], 0)
 # order by dow
 days = ['Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday', 'Saturday', 'Sunday']
 summary['day'] = pd.Categorical(summary['day'], categories=days, ordered=True)
 summary = summary.sort_values(by='day').reset_index(drop=True)
-print(summary)
+
+with open("../results/tables/ecoli_day_of_week_table.md", "w") as file:
+        file.write(summary.to_csv())

### 2024-09-27 code/model_generation/ecoli_scatter.py

@@ -1,6 +1,5 @@
 # Import packages
 import pandas as pd
-import tabulate as tab
 import seaborn as sns
 import matplotlib.pyplot as plt
 import numpy as np
@@ -15,18 +14,18 @@ df['collectiondate'] = pd.to_datetime(df['collectiondate'])
 df['log_ecoli'] = np.log(df['ecoli'])
 
 sns.set(style="whitegrid", context="talk")
-
+plt.figure(figsize=(16, 10))
 # Create a scater plot
-sns.scatterplot(data=df, x='collectiondate', y='log_ecoli', hue='beachname', palette='Set1', s=20, alpha=0.7)
+sns.scatterplot(data=df, x='collectiondate', y='log_ecoli', hue='beachname', palette='Set1', s=100, alpha=0.7)
 
 # Set title and labels
-plt.title('E. coli Levels Over Time by Site and Beach', fontsize=16)
-plt.xlabel('Collection Date', fontsize=14)
-plt.ylabel('E. coli Level (Log(MPN/100mL))', fontsize=14)
+plt.title('E. coli Levels Over Time by Site and Beach', fontsize=20)
+plt.xlabel('Collection Date', fontsize=18)
+plt.ylabel('E. coli Level (Log(MPN/100mL))', fontsize=18)
 plt.grid(True, axis='x')
 # modify x axis labels
-plt.xticks(df['collectiondate'], fontsize=10, rotation=75 )  
-plt.yticks(df['log_ecoli'], fontsize=10)
+plt.xticks(df['collectiondate'], fontsize=14, rotation=75 )  
+plt.yticks(df['log_ecoli'], fontsize=14)
 plt.tight_layout()
 
 plt.gca().xaxis.set_major_locator(MaxNLocator(nbins=17))
@@ -35,5 +34,6 @@ plt.gca().yaxis.set_major_locator(MaxNLocator(nbins=10))
 plt.gca().xaxis.set_major_formatter(plt.matplotlib.dates.DateFormatter('%b %Y'))
 plt.legend(title='Beach Name', loc='upper left', bbox_to_anchor=(1.05, 1))
  
+plt.savefig('../results/figures/ecoli_scatter.png', dpi=300, bbox_inches='tight')
 
 plt.show

### 2024-09-27 code/model_generation/individual_beach_scatter.py

@@ -1,43 +1,0 @@
-
-import pandas as pd
-import matplotlib.pyplot as plt
-import seaborn as sns
-from matplotlib.ticker import MaxNLocator
-
-data = pd.read_csv("../data/cleaned_water_quality_data.csv")
-df = pd.DataFrame(data)
-
-df['collectiondate'] = pd.to_datetime(df['collectiondate'])
-# Separate data by beach
-beach1_df = df[df['beachname'] == 'Sunnyside Beach']
-beach2_df = df[df['beachname'] == 'Marie Curtis Park East Beach']
-
-# Create the plot for Sunnyside Beach
-plt.figure(figsize=(14, 7))
-sns.scatterplot(data=beach1_df, x='collectiondate', y='ecoli', hue='sitename', palette='Set1', s=75, alpha=0.6)
-plt.title('E. coli Levels Over Time - Sunnyside Beach', fontsize=16)
-plt.xlabel('Collection Date', fontsize=14)
-plt.ylabel('E. coli Level', fontsize=14)
-plt.gca().xaxis.set_major_locator(MaxNLocator(nbins=8))
-plt.gca().xaxis.set_major_formatter(plt.matplotlib.dates.DateFormatter('%b %Y'))
-plt.xticks(rotation=45, ha='right')
-  # Log scale for better representation
-plt.grid(visible=True, linestyle='--', linewidth=0.5, alpha=0.7)
-plt.legend(title='Site Name', loc='upper left', bbox_to_anchor=(1.05, 1))
-plt.tight_layout()
-plt.show()
-
-# Create the plot for Marie Curtis Park East Beach
-plt.figure(figsize=(14, 7))
-sns.scatterplot(data=beach2_df, x='collectiondate', y='ecoli', hue='sitename', palette='Set2', s=75, alpha=0.6)
-plt.title('E. coli Levels Over Time - Marie Curtis Park East Beach', fontsize=16)
-plt.xlabel('Collection Date', fontsize=14)
-plt.ylabel('E. coli Level', fontsize=14)
-plt.gca().xaxis.set_major_locator(MaxNLocator(nbins=8))
-plt.gca().xaxis.set_major_formatter(plt.matplotlib.dates.DateFormatter('%b %Y'))
-plt.xticks(rotation=45, ha='right')
-  # Log scale for better representation
-plt.grid(visible=True, linestyle='--', linewidth=0.5, alpha=0.7)
[…]

### 2024-09-27 code/model_generation/monthly_ecoli.py

@@ -11,11 +11,14 @@ df['collectiondate'] = pd.to_datetime(df['collectiondate'], errors='coerce', for
 
 # Add a column for the day of the week
 df['month'] = df['collectiondate'].dt.month_name()
-
+df = df.loc[df['collectiondate'] > '2017-12-31']
 # Group by day of the week and beachname, then calculate average E. coli levels
 average_ecoli = df.groupby(['month', 'beachname'])['ecoli'].mean().reset_index()
 
-# Sort days of the week in the typical order
+# Sort the months
+month_order = ['May', 'June', 'July', 'August', 'September']
+average_ecoli['month'] = pd.Categorical((average_ecoli)['month'], categories=month_order, ordered=True)
+average_ecoli = average_ecoli.sort_values(by='month').reset_index(drop=True)
 
 
 # Create the bar chart
@@ -28,4 +31,7 @@ plt.xticks(rotation=45)
 plt.grid(axis='y', linestyle='--', linewidth=0.5, alpha=0.7)
 plt.legend(title='Beach Name', bbox_to_anchor=(1.05, 1), loc='upper left')
 plt.tight_layout()
+
+plt.savefig('../results/figures/ecoli_monthly.png', dpi=300, bbox_inches='tight')
+
 plt.show()

### 2024-09-27 code/model_generation/monthly_summary.py

@@ -20,7 +20,12 @@ summary = df.groupby('month').agg(
     unsafe_levels=('ecoli', lambda x: (x > 100).sum())
 ).reset_index()
 summary['percentage_unsafe'] = round((summary['unsafe_levels']/summary['n'])*100, 2)
+summary['variance'] = round(summary['variance'], 0)
+summary['average_ecoli']=round(summary['average_ecoli'], 0)
 # order by month
 months = ['May', 'June', 'July', 'August', 'September']
 summary['month'] = pd.Categorical(summary['month'], categories=months, ordered=True)
 summary = summary.sort_values(by='month').reset_index(drop=True)
+
+with open("../results/tables/ecoli_monthly_table.md", "w") as file:
+        file.write(summary.to_csv())

### 2024-09-27 code/model_generation/summary_ecoli.py

@@ -1,0 +1,44 @@
+# Import packages
+import pandas as pd
+import tabulate as tab
+from IPython.display import display
+
+# Import data
+data = pd.read_csv("../data/cleaned_water_quality_data.csv")
+df = pd.DataFrame(data)
+
+df['collectiondate'] = pd.to_datetime(df['collectiondate'])
+
+
+ecoli_summary = df['ecoli'].describe()
+
+# Additional statistics
+ecoli_variance = df['ecoli'].var()
+
+# Create a summary table
+ecoli_stats = pd.DataFrame({
+    'Statistic': [
+        'Count',
+        'Mean',
+        'Standard Deviation',
+        'Min',
+        '25th Percentile (Q1)',
+        'Median (Q2)',
+        '75th Percentile (Q3)',
+        'Max',
+        'Variance'
+    ],
+    'Value': [
+        ecoli_summary['count'],
+        ecoli_summary['mean'],
+        ecoli_summary['std'],
+        ecoli_summary['min'],
+        ecoli_summary['25%'],
+        ecoli_summary['50%'],
+        ecoli_summary['75%'],
+        ecoli_summary['max'],
+        ecoli_variance
+    ]})
+
+with open("../results/tables/ecoli_summary.md", "w") as file:
+        file.write(ecoli_stats.to_csv())

### 2024-09-27 code/model_generation/summary_stats.py

@@ -1,0 +1,31 @@
+# Import packages
+import pandas as pd
+import tabulate as tab
+from IPython.display import display
+
+# Import data
+data = pd.read_csv("../data/cleaned_water_quality_data.csv")
+df = pd.DataFrame(data)
+
+df['collectiondate'] = pd.to_datetime(df['collectiondate'])
+# Summary stats tables
+df = data
+summary_stats = pd.DataFrame({
+    'Metric': [
+        'Data points',
+        'Date Range',
+        'Unique Beach Names',
+        'N0 of Unique Site Names',
+        'No of Unique Locations',
+    ],
+    'Value': [
+        len(df),
+        f"{df['collectiondate'].min()} to {df['collectiondate'].max()}",
+        df['beachname'].nunique(),
+        df['sitename'].nunique(),
+        df.groupby(['longitude', 'latitude']).ngroups
+    ]
+})
+
+with open("../results/tables/data_summary.md", "w") as file:
+        file.write(summary_stats.to_csv())

### 2024-09-27 code/model_generation/summary_tables.py

@@ -1,87 +1,0 @@
-# Import packages
-import pandas as pd
-import tabulate as tab
-import seaborn as sns
-import matplotlib.pyplot as plt
-import numpy as np
-
-# Import data
-data = pd.read_csv("../data/cleaned_water_quality_data.csv")
-df = pd.DataFrame(data)
-
-df['collectiondate'] = pd.to_datetime(df['collectiondate'])
-# Summary stats tables
-def create_counts(data):
-    df = data
-    summary_stats = pd.DataFrame({
-        'Metric': [
-            'Data points',
-            'Date Range',
-            'Unique Beach Names',
-            'N0 of Unique Site Names',
-            'No of Unique Locations',
-        ],
-        'Value': [
-            len(df),
-            f"{df['collectiondate'].min()} to {df['collectiondate'].max()}",
-            df['beachname'].nunique(),
-            df['sitename'].nunique(),
-            df.groupby(['longitude', 'latitude']).ngroups
-        ]
-    })
-    sum_stats = tab.tabulate(summary_stats, headers='keys', tablefmt='github')
-    return sum_stats
-
-
-def create_ecoli_stats(df):
-    ecoli_summary = df['ecoli'].describe()
-
-    # Additional statistics
-    ecoli_variance = df['ecoli'].var()
-
-    # Create a summary table
-    ecoli_stats = pd.DataFrame({
-        'Statistic': [
-            'Count',
-            'Mean',
-            'Standard Deviation',
-            'Min',
-            '25th Percentile (Q1)',
-            'Median (Q2)',
-            '75th Percentile (Q3)',
-            'Max',
-            'Variance'
-        ],
-        'Value': [
-            ecoli_summary['count'],
-            ecoli_summary['mean'],
-            ecoli_summary['std'],
-            ecoli_summary['min'],
-            ecoli_summary['25%'],
-            ecoli_summary['50%'],
-            ecoli_summary['75%'],
-            ecoli_summary['max'],
-            ecoli_variance
-        ]})
-    ecoli_stats = tab.tabulate(ecoli_stats, headers='keys', tablefmt='github')
-    return ecoli_stats
-
-
-
-def main():
-    #Import Data
-    data = pd.read_csv("data/cleaned_water_quality_data.csv")
-    df = pd.DataFrame(data)
-
-    with open("results/tables/summary_table.md", "w") as file:
-        file.write(create_counts(df))
-    
[…]

### 2024-09-27 extra.py

@@ -1,5 +1,0 @@
-import pandas as pd 
-
-data = pd.read_csv("data/cleaned_water_quality_data.csv")
-df = pd.DataFrame(data)
-print(df['loc_type'].unique())

### 2024-09-27 paper/paper.qmd
Paper final draft

@@
 To either close public beaches that exhibit unsafe levels, or continuation of previous action that has led to lower levels.
+
 Despite these public health measures, there is limited understanding of the specific patterns of E.
@@
 By analysing data over several years, this study seeks to identify temporal patterns in contamination, such as variations by day of the week and season, to help answer the critical question: When are beachgoers at the highest risk of swimming in contaminated waters?
+
 Additionally, this analysis explores whether E.
@@
 Or in the worse case, that interventions are needed in order to keep Toronto beaches safe for everyone.
+
 This paper is structured as follows: First, we present the dataset, describing its sources, key variables, and the data cleaning process.
@@
 In this paper, Python [@python] and its packages are utilised as the primary programming language for the following analysis and displaying the data.
-Key packages used include pandas [McKinney, 2010], which facilitated efficient data cleaning, manipulation, and summarization of the dataset, NumPy [@numpy] was employed for numerical operations, Matplotlib [matplotlib] and seaborn [@seaborn] were critical for creating informative visualisations.
+Key packages used include pandas [McKinney, 2010], which facilitated efficient data cleaning, manipulation, and summarization of the dataset, NumPy [@numpy] was employed for numerical operations, Matplotlib [@matplotlib] and seaborn [@seaborn] were critical for creating informative visualisations.
 
 # Data
-
 ## Data Source
@@
 
-(Summary stats)
+\newpage
+```{python}
+#| echo: false
+#| label: fig-0a
+#| fig.pos: "H"
+# Execute the external Python script
+with open(f"../code/model_generation/summary_stats.py") as script_file:
+    exec(script_file.read())
+summary_stats.round(0)
+```
+**Figure 0a** Shows counts of unique variables and the date range of the data
+```{python}
+#| echo: false
+#| label: fig-0b
+#| fig.pos: "H"
+# Execute the external Python script
+with open(f"../code/model_generation/summary_ecoli.py") as script_file:
+    exec(script_file.read())
+ecoli_stats.round(0)
+```
+**Figure 0b** Shows summary statistics for e-coli levels over the whole dataset
 
 ## Measurement
-E-coli is measured in MPN/100ml a measurement that reflects the e coli particles in every 100ml of lake water.
+E-coli is measured in MPN/100 ml a measurement that reflects the e coli particles in every 100 ml of lake water.
 For our data presentation log values are used to give a better visualisation of the data.
@@
 Missing this data may lead to some bias as the time of day could impact the baseline levels of e-coli in the water.
-What this form of measurement does allow us to do is look at the differences in water quality at a date level, a figure() shows there is a good distribution of data from each day of the week.
+What this form of measurement does allow us to do is look at the differences in water quality at a date level, Weekly section shows there is a good distribution of data from each day of the week.
 There are some gaps in the data however, in the monthly data we only find results from the summer months.
@@
 Finally we will look at these distributions over the course of a week to answer the question of when water quality is at its best, and as a lake swimmer when is it safest for me to swim at these beaches.
+For average levels we will focus on data after 2017.
+This is because of the large difference in the variation of the data before and after this year.
+This could be due to a change in testing strategy or something else not represented in the data, making the actual levels reported on either side of this year not comparable.
+The years after 2017 were selected in order to compare more recent data to provide better context.
+The method used in this paper to look at if values are above and below a threshold allow us to compare and use the data as a whole.
 
+\newpage
 # Results
@@
 #| label: fig-1
+#| fig.pos: "H"
 # Execute the external Python script
@@
 
-Figure() displays our data as a whole, we see very widely distributed values of what could be considered high levels of e-coli from 2007 to 2016.
+Figure 1 displays our data as a whole, we see very widely distributed values of what could be considered high levels of e-coli from 2007 to 2016.
 We see the highest density of data at 10 MLN/100ml but on the higher end levels range from 1000 to 9000.
@@
 #| echo: false
+#| fig.pos: "H"
 #| label: fig-2
@@
 
+\newpage
 ## Monthly data
@@
 
-Figure 4 looks at the monthly data at each beach, as mentioned in the data section only the summer months provide data due to the cold weather in Toronto limiting available data in the winter.
+Figure 3 looks at the monthly data at both beaches over the whole dataset, as mentioned in the data section only the summer months provide data due to the cold weather in Toronto limiting available data in the winter.
 Further we see a disproportionate number of data points in May and September, this could make the results here less reliable.
@@
 August and July show extremely high averages, above 500 MPN/100ml, with over 30% of data collected here showing water quality above acceptable levels.
-Looking at figure() these high levels are consistent between both beaches, though we do observe extremely high average levels at Sunnyside.
+Figure 4 shows that these trends are exhibited in both beaches to varying degrees post 2017.
+The high levels in the later 3 months are reflected at both beaches, though we do observe extremely high average levels at Sunnyside compared to Marie Curtis in this period.
+Contrarily at the beginning of the summer there is stellar water quality at sunnyside in May. Marie curtis on the other hand averages above threshold e-coli levels in all months.
 Overall, we see the average level as well as the likelihood of unsafe levels increase as the summer progresses and slightly taper off in september.
-Although the absence of a reasonable amount of data in September could contribute this increasing trend.
+Although the absence of a reasonable amount of data in September could contribute to the ending of this increasing trend.
 
@@
 # Execute the external Python script for weekly summary statistics
-with open("../code/model_generation/weekly_summary.py") as script_file:
+with open("../code/model_generation/dow_summary.py") as script_file:
     exec(script_file.read())
@@
 
-Figure X summarises E.
+Figure 5 summarises E.
 coli data by day of the week, including the total number of samples (n), average E.
@@
 This Is a positive sight as these days, it is assumed, encounter the most activity due to the 5 day work week.
-Figure() gives context to how the two beaches differ in their levels over the course of a week.
-Marie Curtis see's fairly consistent averages across the week, between 300 and 600 MPN/100ml.
-We still observe the lower average observed in the table here however it is far more significant at this location on Saturday compared to Sunday.
-As is clear from the graph we see significant differences in the average in the two locations.
-Sunnyside Beach shows extraordinarily high levels on Monday and Tuesday Compared to Marie Curtis.
-The rest of the weekdays we observe a fairly insignificant difference between the locations average level, but Sunnyside does appear to have a better average quality.
-The weekend we see the inverse to the start of the week, with Sunnyside having very low averages.
-In fact Sunday at Sunnyside is the only category here that has average levels below the safe threshold of 100 MPN/100ml.
+Figure 6 gives context to how the two beaches differ in their levels over the course of a week, in recent years after 2017.
+Marie Curtis see's consistently lower averages over the course of the weekdays, bu boh locations shar equally high levels of e-coli on friday.
+Sunnyside boasts levels over 300MPN/100ml during this period also.
+Highlighting it as having lower water quality consistently.
+Interestingly both locations share very low averages on both weekend days, especially when compared to the other days.
+Comparing this to Figure 5 highlights the improvement the water quality has seen in recent years as both averages now fall very near the observed minimum of 10MPN/100ml.
 
@@
 **Figure 6**: E.
-coli levels by day of the week show significant differences between the two beaches, with Sunnyside exhibiting extraordinarily high levels on Monday and Tuesday compared to Marie Curtis.
+coli levels by day of the week from data after 2017, shows significant differences between the two beaches, with Sunnyside exhibiting extraordinarily high levels on Monday and Tuesday compared to Marie Curtis.
+but both see very low averages on the weekend.
 
@@
 While the data shows a reduction in the variability of E.
-coli levels in recent years, the frequency of unsafe readings remains concerning, particularly during peak summer months.
-The persistent risk highlighted by the bar graph indicates that ongoing contamination remains a public health challenge, despite efforts to improve water quality.
+coli levels in recent years (assuming this chang in variability is not due to some bias), the frequency of unsafe readings remains concerning, particularly during peak summer months.
+The persistent risk highlighted by figure 2 indicates that ongoing contamination remains a public health challenge, despite efforts to improve water quality.Average tests that are above the target level is still over 30% in recent years.
 Policy makers should improve their efforts to reduce the number of days that water quality is unacceptable over the summer months.
-The months that are exhibiting the greatest risk are July and August, as such Policy makers should focus their efforts on this time of the year.
-As these months fall in the middle of the summer the higher levels could be a result of climate or increased usage of the beaches in the warmer weather.
+The latter months are showing the greatest risk as shown in the monthly results section, thus efforts should be made to try and carry the lower levels shown earlier throughout the summer.
 From the perspective of swimmers the results displayed here give insights on how to lower the risks of visiting the beaches when water quality is lower.
 Firstly overall we see that in recent years the extremes are much lower which is good news for lake swimmers in toronto.
-We also see that choosing to swim in the earlier months of the summer appear to bring with them better overall water quality than later months.
+We also see that choosing to swim in the earlier months of the summer appears to bring with them better overall water quality than later months, May is especially low risk.
 Similarly the weekends bear a far lower risk to swimmers, which is fortunate as it is likely that these are the beaches busier times.
-The weekly data does have warning signs also, avoiding Sunnyside during the week, interestingly although this beach has the highest risk days it also pairs that with the safest days in the week.
+The weekly data does have warning signs also, avoiding swimming during the week appears to be the consensus, especially at Sunnyside.
+The weekends, which are assumed to be busier for Toronto's beaches actually exhibit low risk, especially over recent years, this is a very positive sign for public health.
+If this is a result of policy then the data shows that it has been very successful in raising water quality on those days.
+Overall the results here are fairly concerning when it comes to the percentage of results well above and acceptable threshold.
+There are silver linings however, and the insights gathered here can aid swimmers in choosing when and where they can minimise health risks from low water quality.
+The positive data points in the results show where efforts to improve water quality have been successful and where these efforts should be duplicated.
 

### 2024-09-27 paper/paper.qmd
paper final changes

@@
 bibliography: references.bib
+toc: true
+toc-depth: 1
 ---
-
 ```{python}
@@
 
-\newpage
 ```{python}
@@
 
+\newpage
+
 ## Measurement
-E-coli is measured in MPN/100 ml a measurement that reflects the e coli particles in every 100 ml of lake water.
+E-coli is measured in MPN/100ml, a measurement that reflects the e coli particles in every 100 ml of lake water.
 For our data presentation log values are used to give a better visualisation of the data.
@@
 While the data shows a reduction in the variability of E.
-coli levels in recent years (assuming this chang in variability is not due to some bias), the frequency of unsafe readings remains concerning, particularly during peak summer months.
+coli levels in recent years (assuming this change in variability is not due to some bias), the frequency of unsafe readings remains concerning, particularly during peak summer months.
 The persistent risk highlighted by figure 2 indicates that ongoing contamination remains a public health challenge, despite efforts to improve water quality.Average tests that are above the target level is still over 30% in recent years.
